---
title: 'Lab #5'
author: "Kobe Pranivong; illuminobe Yi Hang Khor; bobbykhor"
date: "4/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("dplyr")
library("tidyr")
library("tidyverse")
library("ggplot2")
library("readxl")
```

```{r}
accident <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(accident)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(person)
```

1. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
ggplot(accident, aes(x=DAY_WEEK)) + geom_bar() + scale_x_discrete(name ="Day of the week", 
                    limits=1:7)
```

The last day of the week (day 7) has the greatest number of accidents. In general, the last days of the week (day 6 and 7) experience more accidents than on others.

2. What time of the day do accidents happen (use variable HOUR)?
```{r}
ggplot(accident, aes(x=factor(HOUR))) + geom_bar() + xlim(c(0, 23)) + scale_x_discrete(name ="Time of day (hours)", limits=1:23)
```

Most accidents tend to occur later in the day between 17 hours and 21 hours.

3. What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
```{r}
drunk_count <- accident %>% count(DRUNK_DR > 0)
#drunk_count <- accident %>% filter(DRUNK_DR > 0) %>% summarise(n = n())
drunk_count
```

8769 accidents had at least one drunk driver.

4. In the person table, identify drivers (PER_TYP == 1, see FARS manual) and subset on them.
```{r}
driver <- person %>% filter(PER_TYP == 1)
```

5. Join accident and person table (work out which variable(s) to use)
```{r}
data <- inner_join(driver, accident, by="ST_CASE")
```

6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.
```{r}
data1 <- data %>% select(DAY_WEEK, HOUR.x, SEX) %>%
  filter(DAY_WEEK >= 1 & DAY_WEEK <= 7) %>% 
  filter(HOUR.x >= 0 & HOUR.x <= 23) %>% 
  filter(SEX == 1 | SEX == 2) %>% 
  group_by_all() %>% tally()
data1.1 <- data1 %>% filter(SEX == 1)
data1.2 <- data1 %>% filter(SEX == 2)
ggplot(data1.1, aes(x=factor(DAY_WEEK), y=factor(HOUR.x))) + 
  geom_raster(aes(fill = n), hjust=0.5, vjust=0.5, interpolate=FALSE) + 
  scale_fill_gradientn(colours=c("#FFFFFFFF","#0000FFFF")) +
  scale_x_discrete(name ="Day of the week", limits=1:7) + 
  scale_y_discrete(name ="Time of day (hours)", limits=1:23) +
  labs(title = "Time distribution of accidents (male)")
ggplot(data1.2, aes(x=factor(DAY_WEEK), y=factor(HOUR.x))) + 
  geom_raster(aes(fill = n), hjust=0.5, vjust=0.5, interpolate=FALSE) + 
  scale_fill_gradientn(colours=c("#FFFFFFFF","#FF0000FF")) +
  scale_x_discrete(name ="Day of the week", limits=1:7) + 
  scale_y_discrete(name ="Time of day (hours)", limits=1:23) +
  labs(title = "Time distribution of accidents (female)")
```

7. Now plot a choropleth map of accidents on a county level
```{r}

```

Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication
Use readxl::read_xlsx to read in the GLCs

8. Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.
```{r}

```

9. Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?
```{r}

```